---
title: "M5"
author: "Ryuta Yoshimatsu"
output:
  html_document: 
    number_sections: true
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r packages, message = FALSE}
library(statsr)
library(dplyr)
library(MASS)
library(BAS)
library(ggplot2)
library(devtools)
library(gridExtra)
library(grid)
library(GGally)
library(PreProcess)
library(tidyverse)
library(knitr)

library(TTR)
library(forecast)
library(lubridate)
library(mltools)
library(data.table)
library(ggplotify)
library(gridBase)
library(tsibble)
library(fable)
#library(prophet)
```

# Helper Function

```{r}
# Function to plot Normal QQ
qqplot.data <- function (vec) # argument: vector of numbers
{
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]
  d <- data.frame(resids = vec)
  ggplot(d, aes(sample = resids)) + stat_qq() + geom_abline(slope = slope, intercept = int) + xlab("Theoretical Quantiles") + ylab("Sample Quantiles")
}
```

# First Look at the Data

```{r load, message = FALSE}
calendar <- read.csv("calendar.csv")
sell_price <- read.csv("sell_prices.csv")
evaluation <- read.csv("sales_train_evaluation.csv")
validation <- read.csv("sales_train_validation.csv")
```

```{r}
str(validation[,1:10])
dim(validation)
str(evaluation[,1:10])
dim(evaluation)
```


```{r}
tail(calendar)  # d_1969
dim(validation) # d_1913
dim(evaluation) # d_1941
str(validation[, 1:10]) # first 6 are explanatory variables
```

# Exporatory Analysis

## Total Sales

Create time series object and plot it.

```{r}
total <- validation %>% dplyr::select(-id, -item_id, -dept_id, -cat_id, -store_id, -state_id) %>% summarise(across(everything(), list(sum)))
total <- t(total[,1:ncol(total)])
total.ts <- msts(total, seasonal.periods=c(7, 30.4167, 365.25))

total.eval <- evaluation %>% dplyr::select(-id, -item_id, -dept_id, -cat_id, -store_id, -state_id) %>% summarise(across(everything(), list(sum)))
total.eval <- t(total.eval[,1:ncol(total.eval)])
total.eval.ts <- msts(total.eval, seasonal.periods=c(7, 30.4167, 365.25))
```

Smooth the series to get an insight about the general trend.

```{r, fig.width=15, fig.height=5}
# Moving Average
total.sma <- ts(SMA(total.ts, n=30.4167))
par(mfrow = c(1, 2))
ts.plot(total, total.sma, gpars=list(col = c("black", "red")))
```

Check if there is any significant dependence structures in the series.

```{r, fig.width=5, fig.height=5}
c(total.ts) %>% ggtsdisplay(main='', theme=theme_bw(), lag.max=30)
```

```{r, fig.width=7.5, fig.height=7.5}
total.ts %>% mstl() %>% autoplot()
```

### Holt-Winters & ETS: Exponential Smoothing with Annual Seasonality

Holt-Winters only takes one seasonality in to account, which in this case will be the anual seasonamlity. 

```{r,fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
total.hw <- HoltWinters(total.ts)
total.hw.forecasts <- forecast:::forecast.HoltWinters(total.hw, h=28)
par(mfrow=c(1, 2))
plot(total.hw, xlim=c(4.0, 6.238356))
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)
vp1 <-plotViewport(c(1,1,1,1))
f <- autoplot(total.hw.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.hw.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

Run residual analysis to check the fit.

```{r,fig.width=5, fig.height=5}
# Plot residuals
na.omit(total.hw.forecasts$residuals) %>% ggtsdisplay(lag.max=20, main='Residual', theme=theme_bw())

# Run Ljung-Box test (null hypothesis: there is no non-zero autocorrelation in the in-sample forecast errors)
Box.test(total.hw.forecasts$residuals, lag=20, type="Ljung-Box")

# Plot histogram of residuals
ggplot() + aes(as.numeric(na.omit(total.hw.forecasts$residuals)))+geom_histogram(aes(y=..density..), bins=30, alpha=0.75, color='blue', fill='blue') + geom_density()

# Plot QQ plot
qqplot.data(as.numeric(na.omit(total.hw.forecasts$residuals)))
```

It looks like there are still some statistically significant structures in the residuals. Now we try stlf() from forecast: https://robjhyndman.com/hyndsight/estimation2/#:~:text=HoltWinters()%20is%20using%20heuristic,for%20the%20linear%20additive%20models). 

```{r,fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
total.ets <- stlf(total.ts)
total.ets.forecasts <- forecast(total.ets, h=28)
par(mfrow=c(1, 2))
plot(total.ets.forecasts$x, xlim=c(4.0, 6.238356), col="black")
lines(total.ets.forecasts$fitted, col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)
vp1 <-plotViewport(c(1,1,1,1))
f <- autoplot(total.ets.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.ets.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5, warning=FALSE, message=FALSE}
checkresiduals(total.ets)
sum(total.ets$residuals^2)
```

### Simple ARIMA with Annual Seasonality

Simple ARIMA model only allows one frequency.

```{r}
total.arima <- ts(total, frequency=365.25)
total.simple.arima <- auto.arima(total.arima, D=1, stationary=FALSE, seasonal=TRUE, ic="bic", stepwise=FALSE)
summary(total.simple.arima)
total.simple.arima.forecasts <- forecast(total.simple.arima, h=28)
```

```{r, fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
par(mfrow=c(1, 2))
plot(total.simple.arima$x, xlim=c(4.0, 6.238356), col="black")
lines(fitted(total.simple.arima), col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)        ## I am in the space of the base plot
vp1 <-plotViewport(c(1,1,1,1))  ## Create new vp with margins
f <- autoplot(total.simple.arima.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.simple.arima.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5}
checkresiduals(total.simple.arima)
sum(total.simple.arima$residuals^2)
```

### ARIMA with Weekly, Monthly and Annual Seasonalities and Fourier Terms (i,j,k)

https://robjhyndman.com/hyndsight/forecasting-weekly-data/
https://pkg.robjhyndman.com/forecast/reference/fourier.html
https://otexts.com/fpp2/dhr.html#dhr

The smoothness of the seasonal pattern can be controlled by K (the number of Fourier sin and cos pairs – the seasonal pattern is smoother for smaller values of K).

```{r, fig.width=15, fig.height=5}
start_time <- Sys.time()
bestfit <- list(aicc=Inf)
for(i in 1:3)
  for(j in 1:5)
    for(k in 1:5)
    {
      fit <- auto.arima(total.ts, xreg=fourier(total.ts, K=c(i,j,k)), stationary=FALSE, seasonal=FALSE)
      if(fit$aicc < bestfit$aicc)
        bestfit <- fit
      else break;
    }
end_time <- Sys.time()
end_time - start_time
summary(bestfit)
```

```{r, fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
total.arima.forecasts <- forecast(bestfit, xreg=fourier(total.ts, K=c(2,3,2), h=28))

par(mfrow=c(1, 2))
plot(bestfit$x, xlim=c(4.0, 6.238356), col="black")
lines(fitted(bestfit), col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)        ## I am in the space of the base plot
vp1 <-plotViewport(c(1,1,1,1))  ## Create new vp with margins
f <- autoplot(total.arima.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.arima.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5}
checkresiduals(bestfit)
sum(bestfit$residuals^2)
```

### ARIMA with Multiple Seasonalities (Weekly, Monthly and Annual), Fourier Terms (i,j,k) and Special Days (e.g. Holidays)

https://stackoverflow.com/questions/46873899/weekly-forecasts-with-holidays

```{r, fig.width=15, fig.height=5}
events <- calendar %>% dplyr::select(weekday, event_name_1, snap_CA, snap_TX, snap_WI)
events$saturday <- 0
events$saturday[events$weekday=='Saturday'] <- 1
events$saturday <- as.integer(events$saturday)
events <- events %>% dplyr::select(-weekday)
events <- as.data.frame(model.matrix( ~ . -1, events))
event   <- events[1:1913,]
eventf  <- events[1914:1969,]

# Extend the forecast period to one year
for (i in 1:309)
{
  year_ago <- as.integer(1604+i)
  eventf <- rbind(eventf, event[year_ago,])
}
rownames(eventf) <- 1914:as.integer(1914+nrow(eventf)-1)

h <- nrow(eventf)

fourier_terms <- fourier(total.ts, K=c(2,5,3))
fourier_terms_forecasts <- fourier(total.ts, K=c(2,5,3), h=h)

xreg.fit <- data.matrix(cbind(fourier_terms, event))
xreg.forecasts <- data.matrix(cbind(fourier_terms_forecasts, eventf))

drop_collinear <- rownames(alias(lm(saturday ~ . , data=as.data.frame(xreg.fit)))$Complete)
xreg.fit <- xreg.fit[, !colnames(xreg.fit) %in% drop_collinear]
xreg.forecasts <- xreg.forecasts[, !colnames(xreg.forecasts) %in% drop_collinear]

fit.holidays <- auto.arima(total.ts, xreg=xreg.fit, stationary=FALSE, seasonal=FALSE)
total.arima.holidays.forecasts <- forecast(fit.holidays, xreg=xreg.forecasts, h=h)
```

```{r, fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
par(mfrow=c(1, 2))
plot(fit.holidays$x, xlim=c(4.0, 6.238356), col="black")
lines(fitted(fit.holidays), col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)        ## I am in the space of the base plot
vp1 <-plotViewport(c(1,1,1,1))  ## Create new vp with margins
f <- autoplot(total.arima.holidays.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.arima.holidays.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5}
checkresiduals(fit.holidays)
sum(fit.holidays$residuals^2)
```

### TBATS

https://robjhyndman.com/hyndsight/forecasting-weekly-data/
https://robjhyndman.com/hyndsight/dailydata/

```{r, fig.width=5, fig.height=7.5}
total.tbats <- tbats(total.ts)
plot(total.tbats)
total.tbats.forecasts <- forecast(total.tbats, h=365)
```

```{r, fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
par(mfrow=c(1, 2))
plot(total.ts, xlim=c(4.0, 6.238356), col="black")
lines(total.tbats$fitted.values, col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)        ## I am in the space of the base plot
vp1 <-plotViewport(c(1,1,1,1))  ## Create new vp with margins
f <- autoplot(total.tbats.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.tbats.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5}
checkresiduals(total.tbats)
sum(total.tbats$errors^2)
```

## Neural Network (nnetar)

https://robjhyndman.com/hyndsight/nnetar-prediction-intervals/#:~:text=The%20nnetar%20function%20in%20the,to%20analytically%20derive%20prediction%20intervals.

```{r, fig.width=15, fig.height=5, warning=FALSE, message=FALSE}
total.nn <- nnetar(total.ts)
#total.nn.forecasts <- forecast(total.nn, PI=TRUE, h=28)
total.nn.forecasts <- forecast(total.nn, h=28)

par(mfrow=c(1, 2))
plot(total.nn.forecasts$x, xlim=c(4.0, 6.238356), col="black")
lines(total.nn.forecasts$fitted, col="red")
plot.new()
vps <- baseViewports()
pushViewport(vps$figure)
vp1 <-plotViewport(c(1,1,1,1))
f <- autoplot(total.nn.forecasts) + autolayer(total.eval.ts, series="Data") + autolayer(total.nn.forecasts$mean, series="Forecasts") + xlim(6.238356, 6.312329)
print(f, vp=vp1)
```

```{r, fig.height=5, fig.width=5, warning=FALSE, message=FALSE}
checkresiduals(total.nn)
sum(total.nn$residuals^2)
```

## Prophet

Having trouble loading the library...  

## fable: Multiple 'Independent' Time Series

https://stackoverflow.com/questions/58327544/forecast-multiple-time-series-in-r-using
https://robjhyndman.com/hyndsight/fable/#:~:text=fable%20is%20designed%20for%20tsibble,point%20forecasts%20and%20prediction%20intervals.
https://tidyverts.github.io/tidy-forecasting-principles/index.html

```{r, fig.width=7.5, fig.height=15, warning=FALSE}

# Time series stored in rows
dept_id.df <- validation %>% dplyr::select(-id, -item_id, -cat_id, -store_id, -state_id)
dept_id.df <- dept_id.df %>% group_by(dept_id) %>% summarise(across(everything(), list(sum)))
df <- dept_id.df %>% remove_rownames %>% column_to_rownames(var="dept_id")
dept_id <- as.data.frame(cbind(rownames(df), 1:7))
dept_id <- dept_id %>% rename(Name=V1, Id=V2)
dept_id$Id <- as.integer(dept_id$Id)

# Column names are timestep
colnames(df) <- seq(NCOL(df))

# Convert df to tibble
df <- df %>% as_tibble() 

# Add column 'Id'
df <- df %>% mutate(Id=1:nrow(df))

# Take columns and collapses into key-value pairs
df <- df %>% gather(key="time", value="value", -Id)
df$time <- as.integer(df$time)

# Redefine tsibble object 
df <- df %>% as_tsibble(index=time, key=Id)

# Evaluation data
dept_id.df.eval <- evaluation %>% dplyr::select(-id, -item_id, -cat_id, -store_id, -state_id)
dept_id.df.eval <- dept_id.df.eval %>% group_by(dept_id) %>% summarise(across(everything(), list(sum)))
df.eval <- dept_id.df.eval %>% remove_rownames %>% column_to_rownames(var="dept_id")
colnames(df.eval) <- seq(NCOL(df.eval))
df.eval <- df.eval %>% as_tibble() 
df.eval <- df.eval %>% mutate(Id=1:nrow(df.eval))
df.eval <- df.eval %>% gather(key="time", value="value", -Id)
df.eval$time <- as.integer(df.eval$time)
df.eval <- df.eval %>% as_tsibble(index=time, key=Id)

# Extrenal Regressors
xreg.fit <- data.matrix(event)
xreg.forecasts <- data.matrix(eventf)
drop_collinear <- rownames(alias(lm(event_name_1 ~ . , data=as.data.frame(xreg.fit)))$Complete)
xreg.fit <- xreg.fit[, !colnames(xreg.fit) %in% drop_collinear]
xreg.forecasts <- xreg.forecasts[, !colnames(xreg.forecasts) %in% drop_collinear]

# Fit models
fit <- df %>% model(
  snaive = SNAIVE(value ~ lag(7)),
  ets = ETS(value ~ trend("A") + season("A", period=7) + error("A")),
  arima=ARIMA(value ~ 1 + pdq(p=0:5, d=0:2, q=0:5, p_init=2, q_init=2) + PDQ(P=0, D=0, Q=0) + fourier(7, K=3) + fourier(30.4167, K=5) + fourier(365.25, K=5) + xreg(xreg.fit))
) %>% mutate(mixed=(ets+arima+snaive)/3)

# Compute forecasts
forecasts <- forecast(fit, xreg=xreg.forecasts, h=28)

# Plot forecasts for one series
forecasts %>% autoplot(df.eval) + xlim(1875, 1945)
```

## LGBM

## RNN/LSTM

## Randomly Chosen `id`s

## Vector ARIMA

## Hierarichial Forecast


```{r}
set.seed(12345)
random_ids <- sample(1:nrow(evaluation), 5)
evaluation[c(random_ids), 1:6]
id.df <- t(evaluation[c(random_ids), 7:ncol(evaluation)])
for(i in 1:ncol(id.df))
{
  id.ts <- msts(id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  id.ts.sma <- ts(SMA(id.ts, n=30.4167), frequency=365)
  ts.plot(id.ts, id.ts.sma, gpars=list(col = c("black", "red")))
}
```

## Aggregate Over `item_id` and Select 5 Randomly

```{r, fig.width=7.5, fig.height=5}
set.seed(12345)
item_id.df <- validation %>% dplyr::select(-id, -dept_id, -cat_id, -store_id, -state_id)
item_id.df <- item_id.df %>% group_by(item_id) %>% summarise(across(everything(), list(sum)))
random_item_ids <- sample(1:nrow(item_id.df), 5)
item_id.df <- t(item_id.df[c(random_item_ids), 2:ncol(item_id.df)])
```

```{r}
for(i in 1:ncol(id.df))
{
  item_id.ts <- msts(item_id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  item_id.ts.sma <- ts(SMA(item_id.ts, n=30.4167), frequency=365)
  ts.plot(item_id.ts, item_id.ts.sma, gpars=list(col = c("black", "red")))
}
```

## Aggregate over `dept_id`

```{r}
dept_id.df <- validation %>% dplyr::select(-id, -item_id, -cat_id, -store_id, -state_id)
dept_id.df <- dept_id.df %>% group_by(dept_id) %>% summarise(across(everything(), list(sum)))
dept_id.df <- t(dept_id.df[,2:ncol(dept_id.df)])
```

```{r}
for(i in 1:ncol(dept_id.df))
{
  dept_id.ts <- msts(dept_id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  dept_id.ts.sma <- ts(SMA(dept_id.ts, n=30.4167), frequency=365)
  ts.plot(dept_id.ts, dept_id.ts.sma, gpars=list(col = c("black", "red")))
}
```

## Aggregate Over `cat_id`

```{r}
cat_id.df <- validation %>% dplyr::select(-id, -item_id, -dept_id, -store_id, -state_id)
cat_id.df <- cat_id.df %>% group_by(cat_id) %>% summarise(across(everything(), list(sum)))
cat_id.df <- t(cat_id.df[,2:ncol(cat_id.df)])
```

```{r}
for(i in 1:ncol(cat_id.df))
{
  cat_id.ts <- msts(cat_id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  cat_id.ts.sma <- ts(SMA(cat_id.ts, n=30.4167), frequency=365)
  ts.plot(cat_id.ts, cat_id.ts.sma, gpars=list(col = c("black", "red")))
}
```

## Aggregate Over `store_id`

```{r}
store_id.df <- validation %>% dplyr::select(-id, -item_id, -dept_id, -cat_id, -state_id)
store_id.df <- store_id.df %>% group_by(store_id) %>% summarise(across(everything(), list(sum)))
store_id.df <- t(store_id.df[,2:ncol(store_id.df)])
```

```{r}
for(i in 1:ncol(store_id.df))
{
  store_id.ts <- msts(store_id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  store_id.ts.sma <- ts(SMA(store_id.ts, n=30.4167), frequency=365)
  ts.plot(store_id.ts, store_id.ts.sma, gpars=list(col = c("black", "red")))
}
```

## Aggregate Over `state_id`

```{r}
state_id.df <- validation %>% dplyr::select(-id, -item_id, -dept_id, -cat_id, -store_id)
state_id.df <- state_id.df %>% group_by(state_id) %>% summarise(across(everything(), list(sum)))
state_id.df <- t(state_id.df[,2:ncol(state_id.df)])
```

```{r}
for(i in 1:ncol(state_id.df))
{
  state_id.ts <- msts(state_id.df[,i], seasonal.periods=c(7, 30.4167, 365.25))
  state_id.ts.sma <- ts(SMA(state_id.ts, n=30.4167), frequency=365)
  ts.plot(state_id.ts, state_id.ts.sma, gpars=list(col = c("black", "red")))
}
```
